%%% The main file. It contains definitions of basic parameters and includes all other parts.

%% Settings for single-side (simplex) printing
% Margins: left 40mm, right 25mm, top and bottom 25mm
% (but beware, LaTeX adds 1in implicitly)
\documentclass[12pt,a4paper]{report}
\setlength\textwidth{145mm}
\setlength\textheight{247mm}
\setlength\oddsidemargin{15mm}
\setlength\evensidemargin{15mm}
\setlength\topmargin{0mm}
\setlength\headsep{0mm}
\setlength\headheight{0mm}
% \openright makes the following text appear on a right-hand page
\let\openright=\clearpage

%% Settings for two-sided (duplex) printing
% \documentclass[12pt,a4paper,twoside,openright]{report}
% \setlength\textwidth{145mm}
% \setlength\textheight{247mm}
% \setlength\oddsidemargin{14.2mm}
% \setlength\evensidemargin{0mm}
% \setlength\topmargin{0mm}
% \setlength\headsep{0mm}
% \setlength\headheight{0mm}
% \let\openright=\cleardoublepage

%% Generate PDF/A-2u
\usepackage[a-2u]{pdfx}

%% Character encoding: usually latin2, cp1250 or utf8:
% \usepackage[utf8]{inputenc}

%% Prefer Latin Modern fonts
\usepackage{lmodern}

%% Further useful packages (included in most LaTeX distributions)
\usepackage{amsmath}        % extensions for typesetting of math
\usepackage{amsfonts}       % math fonts
\usepackage{amsthm}         % theorems, definitions, etc.
\usepackage{bbding}         % various symbols (squares, asterisks, scissors, ...)
\usepackage{bm}             % boldface symbols (\bm)
\usepackage{graphicx}       % embedding of pictures
\usepackage{fancyvrb}       % improved verbatim environment
\usepackage{natbib}         % citation style AUTHOR (YEAR), or AUTHOR [NUMBER]
\usepackage[nottoc]{tocbibind} % makes sure that bibliography and the lists
			    % of figures/tables are included in the table
			    % of contents
\usepackage{dcolumn}        % improved alignment of table columns
\usepackage{booktabs}       % improved horizontal lines in tables
\usepackage{paralist}       % improved enumerate and itemize
\usepackage[usenames]{xcolor}  % typesetting in color
\usepackage{float}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{fontspec}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{xunicode} %% loading this first to avoid clash with bidi/arabic
\usepackage{polyglossia}
\usepackage{algorithmic}
% \usepackage{algcompatible}
\usepackage{algorithm}
\usepackage{siunitx}
\usepackage{multicol}
\usepackage{soul}
\setlength{\columnsep}{5mm}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{venndiagram}
\usepackage{colortbl}
\usetikzlibrary{shapes,arrows,fit,calc}

\tikzstyle{decision} = [diamond, draw=black, thick, fill=blue!20, 
    text width=5em, text centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw=black, thick, fill=blue!20, 
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, thick, -latex']


\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\setdefaultlanguage[]{english}
\setotherlanguages{russian,hindi,hebrew,arabic}
\newfontfamily\devanagarifont[Script=Devanagari]{BABEL Unicode}
\newfontfamily\cyrillicfont[Script=Cyrillic]{Charis SIL}
\newfontfamily\hebrewfont[Script=Hebrew]{Noto Sans Hebrew}
\newfontfamily\arabicfont[Script=Arabic]{Noto Naskh Arabic}
% \setkeys{hindi}{numerals=Western}
%%% Basic information on the thesis

% Thesis title in English (exactly as in the formal assignment)
\def\ThesisTitle{Consistency of Linguistic Annotation}

% Author of the thesis
\def\ThesisAuthor{Akshay Aggarwal}

% Year when the thesis is submitted
\def\YearSubmitted{2020}

% Name of the department or institute, where the work was officially assigned
% (according to the Organizational Structure of MFF UK in English,
% or a full name of a department outside MFF)
\def\Department{Institute of Formal and Applied Linguistics}

% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Institute}

% Thesis supervisor: name, surname and titles
\def\Supervisor{RNDr. Daniel Zeman, PhD}
\def\CoSupervisor{Koldo Gojenola, PhD}

% Supervisor's department (again according to Organizational structure of MFF)
\def\SupervisorsDepartment{Institute of Formal and Applied Linguistics}
\def\CoSupervisorsDepartment{Computer Languages and Systems, University of Basque Country (UPV-EHU), Spain}

% Study programme and specialization
\def\StudyProgramme{Computer Science}
\def\StudyBranch{Computational Linguistics}

% An optional dedication: you can thank whomever you wish (your supervisor,
% consultant, a person who lent the software, etc.)
\def\Dedication{%

I would like to express my gratitude to my supervisors, Daniel Zeman from Charles University for his dedicated guidance, his valuable comments as well as his recommendation of the thesis topic; and Koldo Gojenola from UPV/EHU for the much needed push to allow me to keep things structured.

I am very thankful to Prof. Mark\'eta Lopatkov\'a and Prof. Vladislav Kubo\v n from Charles University, Prague and also Dr. Bobbye Pernice from Universit\"at des Saarlandes for their immense help in handling all the difficulties encountered during the entire period of study in LCT programme.

A very special note of thanks is extended to Chiara Alzetta of ItaliaNLP Lab for her constant help in an experiment. From the discussion about the algorithm, to actually running it, the experiment on lisca tools would not have been possible without her valuable inputs.

A very sincere vote of thanks is also extended to Charles University in Prague for the university's provision of computing and storage resources. A very heartfelt thanks is also due, to the brilliant researchers and teachers therein. The thesis would not have been possible without their brilliant shaping of the foundations of the subject within me.

Computational resources were supplied by the Ministry of Education, Youth and Sports of the Czech Republic under the Projects CESNET (Project No. LM2015042) and CERIT-Scientific Cloud (Project No. LM2015085) provided within the program Projects of Large Research, Development and Innovations Infrastructures.

Finally, I would like to thank all the people who supported me during the entire period, my family and friends. Special thanks are due to my friends in Prague for helping me keep my sanity in place.}

% Abstract (recommended length around 80-200 words; this is not a copy of your thesis assignment!)
\def\Abstract{%
This thesis attempts at correction of some identified inconsistencies in different treebanks. The inconsistencies can be related to linguistic constructions, failure of the guidelines of annotation, failure to understand the guidelines on annotator's part, or random errors caused by annotators, among others. We propose a metric to test the POS annotation consistency of different treebanks in the same language, when the annotation guidelines remain the same. We offer solutions to some previously identified inconsistencies in the scope of the Universal Dependencies Project, and check the viability of a proposed inconsistency detection tool in a low-resource setting. The solutions discussed in the thesis are language-neutral, intended to work with multiple languages with efficiency.
}

% 3 to 5 keywords (recommended), each enclosed in curly braces
\def\Keywords{%
{Annotation Consistency} {Annotation Inconsistency} {Error Mining} {Language Independent} {Universal Dependencies} {UD Project} {Syntax} {Morphology}
}

%% The hyperref package for clickable links in PDF and also for storing
%% metadata to PDF (including the table of contents).
%% Most settings are pre-set by the pdfx package.
\hypersetup{unicode}
\hypersetup{breaklinks=true}

% Definitions of macros (see description inside)
\include{macros}

% Title page and various mandatory informational pages
\begin{document}

\include{title}

%%% A page with automatically generated table of contents of the master thesis

\tableofcontents

%%% Abbreviations used in the thesis, if any, including their explanation
%%% In mathematical theses, it could be better to move the list of abbreviations to the beginning of the thesis.
\chapwithtoc{List of Abbreviations}
\begin{itemize}
    \item \textbf{CLAS}- Content-based Labelled Attachment Score
    \item \textbf{GS}- Gold Standard
    \item \textbf{LTR}- Left To Right written order language
    \item \textbf{MWE}- Multi-Word Entity
    \item \textbf{NER}- Named Entity Recognition
    \item \textbf{PCA}- Principal Component Analysis
    \item \textbf{POS}- Part Of Speech
    \item \textbf{RTL}- Right To Left written order language
    \item \textbf{sd}- Standard Deviation
    \item \textbf{SOTA}- State Of The Art
    \item \textbf{TAME}- Time, Aspect, Modality, Evidentiality
    \item \textbf{UD}- Universal Dependencies
\end{itemize}

%%% Each chapter is kept in a separate file
% the pages count refers to experiments completed, in the document


    \include{en/preface}    %     08 pages
    \include{en/problems}    % 06 pages
    \include{en/solutions}   % 04 pages
    \include{en/pos_preface} % 10 pages on Preface to theta_POS and tuning theta_POS for size variation
    \include{en/pos_genre}
    \include{en/pos_conclusion}
    \include{en/conj-head} % 35 pages on conj_head
    \include{en/lisca} % 17 pages on LISCA
    \include{en/AUXvsVERB} % 06 pages on AuxVsVerbs
    \include{en/future} % 06 pages on Future Recommendations
    \include{en/conclusion} % 01 pages on Conclusions
% \include{en/expt01}
% \include{en/expt01-p1}
% \include{en/expt01-final}
% \include{en/expt04} % ?? pages on Non-Projectivity

%%% Bibliography
\include{bibliography}

%%% Figures used in the thesis (consider if this is needed)
\listoffigures

%%% Tables used in the thesis (consider if this is needed)
%%% In mathematical theses, it could be better to move the list of tables to the beginning of the thesis.
\listoftables


%%% Attachments to the master thesis, if any. Each attachment must be
%%% referred to at least once from the text of the thesis. Attachments
%%% are numbered.
%%%
%%% The printed version should preferably contain attachments, which can be
%%% read (additional tables and charts, supplementary text, examples of
%%% program output, etc.). The electronic version is more suited for attachments
%%% which will likely be used in an electronic form rather than read (program
%%% source code, data files, interactive charts, etc.). Electronic attachments
%%% should be uploaded to SIS and optionally also included in the thesis on a~CD/DVD.
%%% Allowed file formats are specified in provision of the rector no. 72/2017.
\appendix
\include{en/appendix} % 8 pages
\openright
\end{document}


% TODO:
% 1. Captions and language names- Use \texttt{} tags